---
title: "Codingchallenge7"
author: "Samit Kafle"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
---
Link to github: [click here](https://github.com/SamitKafle/coding_challenge7.git)

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning = FALSE)
```




**1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor**

```{r}
## Load the required packages
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
## Read the data
STAND <- read.csv("PlantEmergence.csv")

# Convert variables to factors
STAND$Treatment <- as.factor(STAND$Treatment)
STAND$DaysAfterPlanting <- as.factor(STAND$DaysAfterPlanting)
STAND$Rep <- as.factor(STAND$Rep)
```


**2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. **

```{r}
# Fit linear model with interaction
model_interaction <- lm(Emergence ~ Treatment * DaysAfterPlanting, data = STAND)

# Summary and ANOVA
summary(model_interaction)
anova(model_interaction)
```


**3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.**

Based on the results of the linear model in Question 2, the interaction term between Treatment and DaysAfterPlanting is not significant ( p = 1), and therefore, it can be removed
```{r}
## simple model
model_simple <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = STAND)
summary(model_simple)
anova(model_simple)
```

**The intercept of the model is 182.16, which represents the average number of emerged plants under treatment1 at the reference level of Days**

**The coeffiecient for Treatment 2 is -134.53, indicating that Treatment 2 had, on average, 134.53 fewer emerged plants than Treatment 1 (p<0.001), indicating strong negative effect on emergence.**


**4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.**
```{r}
# Estimate least square means
emm <- emmeans(model_simple, ~ Treatment)

# Tukey 
results_lsmeans=cld(emm, alpha = 0.05, details= TRUE)
results_lsmeans
```

*Treatment 2 is significantly different from all other treatments, having the lowest emergence (emmean = 55.6). Treatments 6, 3, and 5 had the highest emergence (emmean ≈ 199–201) and are statistically grouped in group 3, indicating they are significantly different from Treatment 2 and likely Treatment 7.

Treatments 1, 4, 8, and 9 fall in group “23”, meaning they are not significantly different from either group 2 or group 3. This suggests intermediate performance — not as low as Treatment 2 and not as high as the top-performing treatments.*


**5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. **
```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]
  
  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor)))
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters)
  
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

# Example usage
plot_cldbars_onefactor(model_simple, "Treatment")

```

*Treatments sharing the same letter are not significantly different. Treatment 2 (group “c”) had significantly lower emergence than all others. Treatments 3, 5, and 6 (group “a”) had the highest emergence and are not significantly different from each other. Treatments labeled “ab” (1, 4, 8, 9) are not significantly different from either group “a” or “b”*

